---
title: "Greenhouse gas emissions"
author: "Duc-Quang Nguyen"
date: "18 November 2015"
output: html_document
---


```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(countrycode)
library(WDI)
library(magrittr)
```


## Greenhouse gas emissions 

Data downloaded from:

[CAIT](http://www.wri.org/resources/data-sets/cait-historical-emissions-data-countries-us-states-unfccc) CAIT Country GHG Emissions - Last updated: 22 June 2015 CSV


```{r}
### Raw data file
ghg.file <- "data/CAIT_Country_GHG_Emissions_-_csv/CAIT Country GHG Emissions.csv"

### Read data file
ghg.read <- read.csv(ghg.file, skip = 2, 
  check.names = F, stringsAsFactors =  F)
```


### Greenhouse gas emissions (abs, % change, by GDP and by capita)
```{r}
### Total GHG Emissions Excluding Land-Use Change and Forestry (MtCO2e)
data <- ghg.read %>% group_by(Country) %>% 
  summarise( GHG = last(`Total GHG Emissions Excluding Land-Use Change and Forestry (MtCO2e)`)) %>% ungroup()

# get the value in 1990
val1990 <- ghg.read %>% filter(Year == 1990)  %>% group_by(Country) %>% 
  summarise( val1990 = first(`Total GHG Emissions Excluding Land-Use Change and Forestry (MtCO2e)`)) %>% ungroup()

data <- cbind(data, GHG1990 = val1990$val1990)

# data <- cbind(data, shareManufacturing = (ghg.read %>% group_by(Country) %>% 
#   summarise( industrial = last(`Manufacturing/Construction (MtCO2)`)) %>% ungroup() %>% select(industrial) / data$GHG) * 100 )


## get country iso2 country code
data$iso2 <- countrycode(data$Country, "country.name", destination = "iso2c")

# NA iso2
data[which(is.na(data$iso2)),]
# remove NA countries
data <- data[which(!is.na(data$iso2)),]

### get the data year 
ghg.read %>% group_by(Country) %>% 
  summarise( lastT = last(Year))  %>% ungroup() %>% 
  select ( lastT) %>% as.vector() %>% unique() %>% print()

### get the population and GDP-PP
	idw <- c('NY.GDP.MKTP.PP.CD', 'SP.POP.TOTL')

	getwbData <- function (ind, countries = data$iso2, end = 2012) {
		indicator <- WDI(indicator = ind, country = countries, start = 2008, end = end)
		colnames(indicator)[3] <- 'values'
		do.call(rbind, by(indicator, indicator$iso2c, function(ii) {
			rowx <- !is.na(ii[,3])
			if(all(!rowx)) {
				ii[1,]
			} else if (all(rowx)){
				ii[which.max(ii$year),]
			} else {
				ii[which(rowx)[1],]
			}
		}))
	}

	gdp <- getwbData(idw[1], data$iso2)
	pop <- getwbData(idw[2], data$iso2)

	# combine indicators with ghg
  data <- cbind(data, gdp = gdp$values[match(data$iso2, gdp$iso2c)], pop = pop$values[match(data$iso2, pop$iso2c)])
	
  # remove all NA
	data <- data[-unique(which(is.na(data), T)[,1]),]
	
	iso2.subset <- c('AT', 'AU', 'KW', 'AE', 'CA', 'US', 'BR', 'RU', 'CN', 'FR', 'IT', 'DE', 'GB', 'SP', 'PT', 'JP', 'TR', 'IN', 'SA', 'MA', 'MX', 'ID', 'ZA', 'EG', 'CH')
	
	df <- data %>% filter (iso2 %in% iso2.subset)
	df$diffGHG  <- round((df$GHG - df$GHG1990) / df$GHG1990, 2) * 100
	df$GHGperCapita <- round(df$GHG * 10^6 / df$pop,1)
	df$GHGperGDP_PPP <- round(df$GHG * 10^6 / df$gdp * 10^6)
	
	
	## reformat for datawrapper
	dw <- df %>% select (Country, iso2, GHG, diffGHG, GHGperCapita, GHGperGDP_PPP)
	dw %<>% rename (
	  `Total greenhouse gas emissions (megatonne CO2, 2012)` = GHG, 
	  `% change in greenhouse gas emissions 1990-2012` = diffGHG,
	  `Greenhouse gas emissions per capita (tonne CO2 per capita, 2012)` = GHGperCapita,
		`Greenhouse gas emissions per GDP (tonne CO2 per million USD GDP, 2012)*` = GHGperGDP_PPP
	  )
	
	write.csv(dw, file = "output/ghgEmissions.csv", row.names = F)
	
```

### Historical carbon emissions

Data downloaded from http://www.wri.org/resources/data-sets/cait-historical-emissions-data-countries-us-states-unfccc

```{r}
### Raw data files
hce.file <- "data/CAIT_Country_GHG_Emissions_-_csv/CAIT Country CO2 Emissions.csv"

### Read data files
hce.read <- read.csv(hce.file, skip = 1, check.names = F, stringsAsFactors =  F)
hce.read$iso2 <- countrycode(hce.read$Country, "country.name", destination = "iso2c")

# ensure the world value is the same of the sum of non-aggregated countres
test <- cbind(hce.read %>% filter(Country == "World"), testSum = 
  hce.read %>% filter(!is.na(iso2)) %>% group_by(Year) %>% 
  summarise(tot = sum(`Total CO2 Emissions Excluding Land-Use Change and Forestry (MtCO2)`, na.rm = T)) %>% ungroup() %>% select(tot))
## IT is NOT the same!

hce.read %>% group_by(Country) %>% 
  summarise(tot = sum(`Total CO2 Emissions Excluding Land-Use Change and Forestry (MtCO2)`, na.rm = T)) %>%
  arrange(desc(tot)) 

country.sub <- c('United States', 'European Union (28)', 'China', 'Russian Federation', 'Japan', 'India', 'Switzerland')


hce.sub <- hce.read %>% filter(Country %in% c("World", country.sub))
# compute "others" values
world <- hce.sub %>% filter(Country == "World") %>% group_by(Year) %>% summarise(world = `Total CO2 Emissions Excluding Land-Use Change and Forestry (MtCO2)`)
sub.tot <- hce.sub %>% filter(Country != "World") %>% group_by(Year) %>% 
  summarise(tot = sum(`Total CO2 Emissions Excluding Land-Use Change and Forestry (MtCO2)`, na.rm = T))
world$world <- world$world - sub.tot$tot
world$world <- ifelse(world$world < 0, 0, world$world )

# reassign others value!
hce.sub[which(hce.sub$Country == "World"), "Total CO2 Emissions Excluding Land-Use Change and Forestry (MtCO2)"]  <- world$world
hce.sub[which(hce.sub$Country == "World"), 'Country'] <- "others"


# replace NA emissions by 0
hce.sub[is.na(hce.sub[,3]),3] <- 0

# compute the cumulative sum
hce.sub %<>% group_by(Country) %>% mutate(cumulative = cumsum(`Total CO2 Emissions Excluding Land-Use Change and Forestry (MtCO2)`)) %>% ungroup()

# compute the cumulative share
hce.sub %<>%  group_by(Year) %>% mutate(cumulativeShare = (cumulative / sum(cumulative)) * 100)  %>% ungroup()


write.csv(hce.sub, file = "output/historicalEmissions.csv", row.names = F)

